---
title: "Prediction Assignment"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(caret)
library(gbm)
library(elasticnet)
library(mlbench)
library(parallel)
library(doParallel)
```

```{r}
cluster <- makeCluster(detectCores()-1)
registerDoParallel(cluster)

fitControl <- trainControl(method="cv", number=5, allowParallel=TRUE)
```


#### Load the training and test data
```{r}
training_data_raw <- read.csv("pml-training.csv", na.strings = "NA")
testing_data_raw <- read.csv("pml-testing.csv")
```

#### Cleaning the data



Columns have either no missing values, or at least 98% missing values.
Columns containing almost exclusively missing values are removed from the data set.


Variables needed for prediction:

readouts of sensors

- 4 sensors at arm, forearm, dumbell and belt

- each sensor has x-y-z, for gyros, accel and magnet

- combined for roll, pitch, yaw and total accel

- other vars to be removed

- roll x,y,z of Arm, Forearm, Belt, Dumbell

- gyros

- accel

- magnet

```{r}
test <- testing_data_raw[colSums(!is.na(testing_data_raw))>0]
training <- training_data_raw[colSums(!is.na(testing_data_raw))>0]
```

59 variables remain to predict the movement class, after removing NA values from the test and training data set.


Remove sample id, variables for recorded time and user name from training data. (columns 1-6)
```{r}
training <- training[-c(1:6)]
```


#### Exploratory Data Analysis

Test data consists of 20 problem id's, which are a placeholder for the actual class of movement (A,B,C,D,E)


```{r}
unique(training$classe)
```



In order to cross validate our trained model, we need to split the training data set. 
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.7, list=FALSE)

training <- training[inTrain,]

testing <- training[-inTrain,]

```


### Random Forest

Fitting the model
```{r cache=TRUE}
set.seed(121)
fit_rf <- train(classe~., method="rf", training, trControl = fitControl)
```

Prediction and test for accuracy
```{r}
pred_rf <- predict(fit_rf, testing)
confusionMatrix(pred_rf, testing$classe)
```

Accuracy of .9981 for the model on the split data from the original training data set.

Now we can predict the variables from the original test set and combine the results in for each ID in a new data frame.
```{r}
pred_final <- predict(fit_rf, test)

#creating data frame for results

results <- cbind(test$X, data.frame(pred_final))
results
```
We achieved a result of 20/20 correct predictions for the test set.

```{r}
stopCluster(cluster)
registerDoSEQ()
```
